---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
---

# Univariate analysis
### Get demographic data from community report
The community data come from seattle government open data portal, which contain census results of 52 communities in seattle in 2013.
```{r }
community=read.csv('https://github.com/chjy0205/data/raw/master/Community%20Reporting%20Areas.csv',stringsAsFactors = F)
```
### Verify data types
```{r}
str(community)
```
### Centrality
It is noticeable that there is a huge gap between minimum median household income and maximum one. My following analysis explores the potential causes of this situation.
```{r}
summary(community$Median.household.income..dollars.)
```
### Dispersion
```{r}
library(moments)
sd(community$Median.household.income..dollars.,na.rm= T)
sd(community$Median.household.income..dollars.,na.rm =T)/mean(community$Median.household.income..dollars.,na.rm =T)
```
### Skewness
```{r}
skewness(community$Median.household.income..dollars.,na.rm=T)
```
### Kurtosis
```{r}
kurtosis(community$Median.household.income..dollars.,na.rm = T)

data=community[is.finite(community$Median.household.income..dollars.),]
var=data$Median.household.income..dollars.
mnVar=mean(var,na.rm = T)
sdVar=sd(var,na.rm = T)

library(ggplot2)

base=ggplot(data, aes(x=var))
hist = base + geom_histogram(fill="green", 
                             color='grey',
                          aes(y=..density..))
histAndNormal = hist + stat_function(fun=dnorm,
                                     color="red",
                                     args=list(mean=mnVar,sd=sdVar))

histAndNormal
```

```{r}
mdVar=median(var)
histAndNormal + geom_vline(aes(xintercept = mnVar,colour="mean"),
                           show.legend  = TRUE) + 
                geom_vline(aes(xintercept = mdVar,colour="median"),
                           show.legend  = TRUE) + 
                scale_color_manual(name = "centrals", 
                                   values = c(median = "blue", mean = "red"))
```

# Clustering
I try to cluster the communities based on their median income, degree and percentage of people of color.
From the result we can see that there are three main groups.
```{r}
length(community$Community.Reporting.Area.Name)==length(unique(community$Community.Reporting.Area.Name))

row.names(community)=community$Community.Reporting.Area.Name

str(community[,c(11,15,21)])

library(cluster)
incoSimi=daisy(community[,c(11,15,21)],metric = "euclidean",
               stand=TRUE)
incoClus=hclust(incoSimi,method = 'average')
plot(incoClus,cex=0.5)

```


# Linear Regression
This part tests the correlation among median income(Y variable), degree and population of color(X variables).

### Check the data
```{r}
hist(community$Median.household.income..dollars.)
```

### Verify that each variable in X has a linear relationship with Y.
```{r}
explanans=names(community)[c(11,15)]
for (x in explanans){
    par=cbind(community[,x],community$Median.household.income..dollars.)
    p=cor.test(community[,x],community$Median.household.income..dollars.)
    print(paste("Pearson:",p$estimate," - Is significative?",p$p.value<0.05))
}
```
### Check the correlation between each pair of X variables.
```{r}
cor(community[explanans])
```
### Set the row names with the community names.
```{r}
row.names(community)=community$Community.Reporting.Area.Name
```
### Request a linear regression model and check the detailed result.
```{r}
LinRegINCOME=lm(community$Median.household.income..dollars.~ Percent.of.Population.of.Color + Percent.Bachelor.degree.or.higher, data = community)

summary(LinRegINCOME)
```
### Regression of income and percent population of color
The results above shows that the coefficient of two X variables are not significant enough because there is strong correlations between degree and population of color percent. Thus, I run regression of single X variable.
```{r}
LinRegINCOME1=lm(community$Median.household.income..dollars.~Percent.of.Population.of.Color, data=community)
summary(LinRegINCOME1)
```
This result is statistically significant which indicates that holding all other factors constant, one percent increase in percent of population of color is associated with $642.2 decrease in median household income.

### Regression of income and percent bachelor degree or higher
```{r}
LinRegINCOME2=lm(community$Median.household.income..dollars.~Percent.Bachelor.degree.or.higher, data=community)
summary(LinRegINCOME2)
```
This result is statistically significant which indicates that holding all other factors constant, one percent increase of population with bachelor degree or higher is associated with $851.3 increase in median household income.

### Regression of degree and population of color
```{r}
LinRegINCOME3=lm(community$Percent.Bachelor.degree.or.higher~Percent.of.Population.of.Color, data=community)
summary(LinRegINCOME3)
```
This result is statistically significant which indicates that holding all other factors constant, one percent increase of population of color is associated with 0.66% decrease in population with bachelor degree or higher.

# Map
It is more obvious to see the distribution of three variables on the map.
```{r}
compressedMap="https://github.com/chjy0205/PROJECT/raw/master/Neighborhoods.zip"
library(utils)
temp=tempfile()
download.file(compressedMap, temp)
unzip(temp)

(maps=list.files(pattern = 'shp'))

library(rgdal)
SeattleMap <- rgdal::readOGR("Neighborhoods.shp",stringsAsFactors=F)
```
### Merge two tables
```{r}
str(SeattleMap$S_HOOD)
str(community$Community.Reporting.Area.Name)

MSeattleMap=merge(SeattleMap,community, by.x='S_HOOD', by.y='Community.Reporting.Area.Name',all.x=F)
nrow(MSeattleMap)
```

```{r}
plot(SeattleMap, col='black')
plot(MSeattleMap, col='pink',border='grey',add=T)
```

### Define the input
```{r}
library(RColorBrewer)
library(classInt)
varToPlot=MSeattleMap$Percent.of.Population.of.Color
varToPlot2=MSeattleMap$Percent.Bachelor.degree.or.higher
varToPlot3=MSeattleMap$Median.household.income..dollars.
```
### Get colors and intervals
```{r}
numberOfClasses = 3
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPlot, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)
colorPallette <- findColours(intervals, colors)
```
### Plot variable1
```{r}
legendText="Population of color in Percent"
shrinkLegend=0.6
title="Population of color in Percent in Seattle 2013"

plot(SeattleMap,col='grey',border='white',main=title)
plot(MSeattleMap, col = colorPallette,border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```

Population of color are mainly distributed in the south part of Seattle.

### Plot variable2
```{r}

numberOfClasses = 3
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPlot2, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)
colorPallette <- findColours(intervals, colors)


legendText="Bachelor Degree or higher in Percent"
shrinkLegend=0.6
title="Population with bachelor degree or higher in Seattle 2013"

plot(SeattleMap,col='grey',border='white',main=title)
plot(MSeattleMap, col = colorPallette,border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```

Communities in the north Seattle have higher ratio of residents holding bachelor degree or higher. Compared with last plot, south Seattle communities have much lower ratios of bachelor degree or higher residents.

### Plot variable3
```{r}
numberOfClasses = 3
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPlot3, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)
colorPallette <- findColours(intervals, colors)


legendText="Median household income in dollar"
shrinkLegend=0.6
title="Median income in Seattle 2013"

plot(SeattleMap,col='grey',border='white',main=title)
plot(MSeattleMap, col = colorPallette,border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```

The median household income roughly obey the trend showed in last two plots. Higher median household income appears more in north Seattle. South Seattle communities often have lower median household income.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
